Are rankings of financial analysts useful to investors?

### 2015-03-11 doc/paper/chapter-1.Rnw
clean workspace for the true, naive, default cases. Predictions are in Black-Lit

% \documentclass{article}
% \usepackage{fancyhdr}
% \usepackage{time}
% \usepackage{authblk}
% \usepackage{amsmath, amssymb}
% \usepackage{caption}
% \captionsetup[table]{textfont={large},labelfont={normalsize},labelsep=newline,justification=justified,singlelinecheck=true, skip=0pt}
% \captionsetup[figure]{textfont={large},labelfont={normalsize},labelsep=newline,justification=justified,singlelinecheck=true, skip=0pt}
% %\usepackage{fullpage}
% %\usepackage{rotating}
% %\usepackage{xcolor}
% \usepackage[hyphens]{url}
% \usepackage[colorlinks=true,urlcolor=blue,citecolor=black]{hyperref}
% %\usepackage{breakurl}
% \usepackage{times}
% \usepackage{setspace}
% \usepackage{graphicx}
% \usepackage{array}
% \usepackage{tabularx}
% \usepackage{booktabs}
% \usepackage{multirow}
% %\onehalfspacing
% \doublespace
% \usepackage[longnamesfirst]{natbib}
% %\usepackage[natbibapa,longnamesfist]{apacite}
% 
% \DeclareMathOperator*{\argmax}{arg\,max}
% %\newcommand{\r}{\Sexpr{}}
% \DeclareMathOperator{\rank}{rank}
% \DeclareMathOperator*{\median}{median}
% %\newcommand{\theHalgorithm}{\arabic{algorithm}}
% \newcommand{\tr}{\textit{true}}
% \newcommand{\naive}{\textit{recent}}
% \newcommand{\default}{\textit{all-time}}
% \newcommand{\market}{\textit{market}}
% \newcolumntype{Y}{>{\raggedleft\arraybackslash}X}% raggedleft column X
% %\newcommand{\raw}{\textit{raw}}
% %\newcommand{\diff}{\textit{diff}}
% %\newcommand{\random}{\textit{random}}
% %\newcommand{\rollsd}{\textit{roll.sd}}
% 
% 
% \newcommand{\ra}[1]{\renewcommand{\arraystretch}{#1}}
% 
% %-------------------------------------------------------------------------
% % take the % away on next line to produce the final camera-ready version
% % Be sure to remove \thispagestyle{fancy} as well after the \maketitle.
% %\pagestyle{empty}
% \pagestyle{fancy}
%  
% \newcommand\myTime{\now}
% \fancyhead{}
% \fancyhead[CO, CE]{\texttt{-Draft-}}
% \fancyhead[RO, RE]{\texttt{\today, \myTime}}
% \setlength{\headheight}{2\baselineskip}
% \renewcommand{\headrulewidth}{0pt}
% %-------------------------------------------------------------------------
% 
% 
% \begin{document}

@@

<<chp1-load-data,echo=F,warning=FALSE,message=FALSE>>=
strategies <- c('true','recent','all-time')
setwd('~/Dropbox/workspace/Projects/Black-Litterman/BL-strategies')
library(reshape2)
library(descr)
library(data.table)
library(ggplot2)
library(stringr)
library(zoo)
library(xtable)
library("PerformanceAnalytics")
load('cache/q.data.RData')
load('cache/final.bl.RData')
load('cache/market.set.RData')
load('cache/example.dt.RData')
source('lib/BL-functions.R')
#library(ProjectTemplate)
#load.project()
###ADD NAs for the case of no ranking ( if NA, no raking)
trunk.percent <- 0.05
stocks <- unique(market.set[,Stock])
#q.data[,quantile(b.view,c(trunk.percent,1-trunk.percent),na.rm=T)]
q.data.full <- setkey(q.data[,':='(year=format(s.Date,'%Y'))],Stock)

@@

brok.full.stat.pt <- rbind(setkey(na.omit(core.dt)[,.N,by=.(Stock,q.id,year)][,list(min=min(N),mean=mean(N),median=as.character(median(N)),max=max(N),st.dev=sd(N)),by=year],year),setnames(cbind('Total',na.omit(core.dt)[,.N,by=.(Stock,q.id,year)][,list(min=min(N),mean=mean(N),median=as.character(median(N)),max=max(N),st.dev=sd(N))]),1,'year'))

#Views stat

stat.nr <- rbind(na.omit(nr.views)[,list(mean=mean(View),median=median(View),st.dev=sd(View),Stocks=length(unique(Stock))),by=.(year,Method)],setnames(cbind('Total',na.omit(nr.views)[,list(mean=mean(View),median=median(View),st.dev=sd(View),Stocks=length(unique(Stock))),by=.(Method)]),1,'year'))[,Strategy:='CONS']

stat.pt.view <- rbind(pt.rank.views[View!=0,list(mean=mean(View),median=median(View),st.dev=sd(View),Stocks=length(unique(Stock))),by=.(year,Method)],setnames(cbind('Total',pt.rank.views[View!=0,list(mean=mean(View),median=median(View),st.dev=sd(View),Stocks=length(unique(Stock))),by=.(Method)]),1,'year'))[,Strategy:='PT']

stat.eps.view <- rbind(eps.rank.views[View!=0,list(mean=mean(View),median=median(View),st.dev=sd(View),Stocks=length(unique(Stock))),by=.(year,Method)],setnames(cbind('Total',eps.rank.views[View!=0,list(mean=mean(View),median=median(View),st.dev=sd(View),Stocks=length(unique(Stock))),by=.(Method)]),1,'year'))[,Strategy:='EPS']

@@

brok.full.stat.eps <- rbind(setkey(na.omit(complete.dt)[,.N,by=.(Stock,q.id,year)][,list(min=min(N),mean=mean(N),median=as.character(median(N)),max=max(N),st.dev=sd(N)),by=year],year),setnames(cbind('Total',na.omit(complete.dt)[,.N,by=.(Stock,q.id,year)][,list(min=min(N),mean=mean(N),median=as.character(median(N)),max=max(N),st.dev=sd(N))]),1,'year'))

@@

% \title{Are rankings of financial analysts useful to investors?}
%  \author[1,2]{ Artur Aiguzhinov (\href{mailto:artur.aiguzhinov@inescporto.pt}{artur.aiguzhinov@inescporto.pt})}
%  \author[1]{ Ana Paula Serra (\href{mailto:aserra@fep.up.pt}{aserra@fep.up.pt})}
%  \author[2,4]{Carlos Soares (\href{mailto:csoares@fe.up.pt}{csoares@fe.up.pt})}
% 
% \affil[1]{FEP \& CEF.UP, University of Porto}
% \affil[2]{INESC TEC}
% \affil[4]{FEUP, University of Porto}
% 
% 
% \maketitle
% % You need this here, or else the first page won't have a header.
% \thispagestyle{fancy}
% 
% \begin{abstract}
% Some institutions issue rankings of financial analysts. Given that these rankings are \textit{ex-post} they may not be able useful to investors. In this paper  we show that trading strategies based on past rankings outperform a passive strategy.  In addition, we report that analysts issuing accurate price targets are more favorable to the investors  to follow rather than analysts issuing accurate EPS forecasts. In addition, we report that the performance of trade strategies based upon price target accuracy is superior one  based upon EPS forecast accuracy
% %Finally, we show that a strategy based on the perfect foresight of analysts' performance gains the highest cumulative return.
% \end{abstract}

\chapter{Are rankings of financial analysts useful to investors?}
\section{Introduction}
\label{sec:intro}

@@

The paper is organized as follows: section \ref{sec:ranking} provides motivation to use rankings of financial analysts; section \ref{sec:trading} outlines our proposed  trading strategies; section \ref{sec:rankings} describes the sample and presents some preliminary results; section \ref{sec:results} presents and discusses the results; and section \ref{sec:conclusion} concludes.

\section{Industry Rankings of Financial Analysts}
\label{sec:ranking}

@@

%In addition, for each strategy, we have three timing scenarios of when we know the analysts' information: we use the \tr{} timing for the case when the information is known at time $t$ (a hypothetical case); we use the \naive{} timing for the $t-1$ case, and we use the \default{} timing \citep{aiguzhinov2010} for the $t=1 \ldots t-1$ case.

@@

\section{Trading Strategies}
%- BL model
%- Information sets
\label{sec:trading}
Our trading strategy use the framework for  active portfolio management proposed by \cite{black1992}.  The model incorporates ``views" in a CAPM framework, forming optimal portfolios in a mean-variance optimization setting. ``Views" are expectations on individual stocks future performance by investors/analysts.

@@

We define a trading strategy as follows (figure \ref{fig:bl}):
\begin{enumerate}
\item  At the beginning of quarter $t$ for each stock $s$,   we define $Q$ and $\Omega$ (see \ref{def-q} and \ref{def-omega} below);

@@

\subsection{Defining $Q$}
\label{def-q}

For the consensus strategy, we use median of expected returns for a particular stock $s$:
\begin{equation}
\label{consq}
Q_{cons,s}= \median \left\{r_{j,s}\right\}
%\frac{1}{N} \sum_{j=1}^{N} r_{j,s}
\end{equation}
%$N$ is the number of analysts with a valid TP report and
where $r_{j,s}=TP_{j,s}/P_{s}-1$  is analyst's $j$ expected return computed using the analyst price target $TP_{j,s}$ and stock price $P_{s}$\footnote{Consistent with the literature, we use stock price 3 days \emph{ex-ante} the TP announcement. This is done to avoid any information leakage around new TP announcement day \citep{bonini2010}}.

@@

\begin{equation}
\label{eq:weight}
w_{j,s}=1-\frac{rank_{j,s}-\min{ \{rank \} }}{\max{\{rank \}}}
\end{equation}

The expected rank-weighted return is thus:
\begin{equation}
\label{rankq}
Q_{rank,s}=\frac{\sum_{j=1}^{N} (w_{j,s} \times r_{j,s})}{\sum_{j=1}^{N} w_{j,s}}
\end{equation}
$N$ is the number of analysts.

@@

\begin{equation}
\label{dfe}
FE_{i,s}^{TP}=|{P_{s}-TP_{j,s}}|
\end{equation}
The PMAFE is given as:
\begin{equation}
\label{tp:pmafe}
PMAFE_{j,s}^{TP}=\frac{FE_{i,s}^{TP}}{\overline{FE_{s}^{TP}}}
\end{equation}
where $\overline{{FE}_{t,s}^{TP}}$ is the average over quarter forecasting error. The target price is fixed over the quarter unless it gets revised.

The rank  that enters equation (\ref{eq:weight}) is average analyst's $PMAFE^{TP}$ per quarter:
\begin{equation}
\overline{PMAFE_{j,s}^{TP}}=\frac{1}{T} \sum_{t=1}^{T} PMAFE_{j,t,s}^{TP}
\end{equation}

\begin{equation}
\label{tp:rank}
rank_{j,s}=\rank_{j=1}^{N} \left\{ \overline{PMAFE_{j,s}^{TP}} \right\}
\end{equation}
where $T$ is the number of trading days in a quarter. Figure \ref{fig:example} shows an example.

\subsubsection{EPS ranking} 
To compute the EPS rankings, we apply the same procedure as above:
\begin{equation}
FE_{j,s}^{EPS}=|{ACT_{s}-PRED_{j,s}}|
\end{equation}
\begin{equation}
PMAFE_{j,s}^{EPS}= \frac{FE_{j,s}^{EPS}}{\overline{FE_{s}^{EPS}}}
\end{equation}
\begin{equation}
\label{eps:rank}
rank_{j,s}=\rank_{j=1}^{N} \left\{ PMAFE_{j,s}^{EPS} \right\} 
\end{equation}
where $ACT_{s}$ and $PRED_{j,s}$ are the actual and  analyst $j$ forecast EPS respectively for each stock $s$.

\subsection{Defining the confidence of expected returns $\Omega$}
\label{def-omega}
The confidence of $Q$ are given by the coefficient of variation of forecasting errors:

\begin{equation}
CV = \frac{\sigma (FE)}{\overline{FE}}
\end{equation}
where $\sigma$ and $\overline{FE}$ are the standard deviation and the mean of the analysts' forecasting error for either TP or EPS. A low value of $CV$ reflects consensual estimates of either future prices or EPS.

@@

In addition to these sets,  we also create a hypothetical scenario that assumes we know the future information about the analyst' accuracy performance which would only be available at the end of $t$. In this setting, we are assuming that we could  perfectly predict the true analysts' ranking in advance. This represents the perfect foresight strategy; therefore, it serves as a reference point to the other trading strategies. We call this approach as the \tr{} set.

We formalize these sets as follows. 
\begin{itemize}
\item  the \tr{} %-- a perfect foresight information:
\begin{equation}
\label{q:true}
\widehat{Q_{.,t}}=Q_{.,t}
\end{equation}

\item  the \naive{} % -- $t-$ information:
\begin{equation}
\label{q:naive}
\widehat{Q_{.,t}}=Q_{.,t-1}
\end{equation}

\item  the \default{} %-- the entire history of analysts performance:
%\begin{itemize}
%\item ranking based views:
\begin{equation}
\label{q:default}
\widehat{Q_{.,t}} = \frac{1}{T} \sum_{t=1}^{T} Q_{.,t}
\end{equation}
%\item consensus based views:
%\begin{equation}
%\label{default.cons}
%\widehat{consQ_{t,s}} = \frac{1}{T} \sum_{T=1}^{T-1} consQ_{t,s}
%\end{equation}
%\end{itemize}
\end{itemize}
where the $.$ subscript corresponds to either $Q_{cons,s}$ (eq. \ref{consq}) or $Q_{rank,s}$ (eq. \ref{rankq}).

\section{Data and preliminary results}
\label{sec:rankings}

@@

Consistent with other studies on analysts' expected returns that work with price targets  \citep{bradshaw2002,brav2003,da2011}, we truncate the sample of $TP/P-1$ at the \Sexpr{ trunk.percent *100}\textsuperscript{th} percentile (values below \Sexpr{round(q.data[,min(trunk.view,na.rm=T)],3)}) and at the \Sexpr{(1-trunk.percent)*100}\textsuperscript{th} percentile (values above \Sexpr{ round(q.data[,max(trunk.view,na.rm=T)],3)}). This is done due to occurrence of the extreme values. Most of these extreme values are driven by misalignment errors found on I/B/E/S data\footnote{We found some differences between the  DataStream and I/B/E/S the databases. In some cases the stock-splits and the dividends were not properly adjusted.}. After trimming, the number of observations ($\mathrm{Stock} \times \mathrm{Broker} \times  \mathrm{Quarter}$) is reduced  from \Sexpr{ prettyNum(q.data[,.N], big.mark=' ')} to \Sexpr{ prettyNum(q.data[!is.na(trunk.view),.N],big.mark=' ')}.

The initial file of quarterly EPS forecast consists of  \Sexpr{eps.dt[,.N,by=Broker][,.N]} brokers covering \Sexpr{eps.dt[,.N,by=Stock][,.N]} stocks. To implement the EPS ranking, we require that a stock had at least three brokers per quarter. In addition, we require that a broker has to be active in covering a particular stock for at least 3 years (12 quarters). Our final sample of EPS forecasts consist of  \Sexpr{complete.dt[,.N,by=Broker][,.N]} brokers covering \Sexpr{complete.dt[,.N,by=Stock][,.N]} stocks. The total number of observation is \Sexpr{prettyNum(complete.dt[,.N],big.mar=' ')}.

Table (\ref{tab:ret-stat}) shows the distribution of the final sample of target price and EPS data. Panel A shows the number of quarterly target prices per stock. For the sample period (1999-2009), we report \Sexpr{prettyNum(core.dt[,.N,by=list(Stock,q.id)][,.N],big.mark=' ')} stock-quarter observations. Each stock had on average of \Sexpr{round(core.dt[,.N,by=list(Stock,q.id)][,mean(N)],2)} quarterly target price reports. Panel B  presents similar statistics of the number of EPS forecasts. In total, the EPS forecast sample has \Sexpr{prettyNum(complete.dt[,.N,by=list(Stock,q.id)][,.N],big.mark=' ')} stock-quarter observations. The average number of quarterly forecasts is \Sexpr{round(complete.dt[,.N,by=list(Stock,q.id)][,mean(N)],2)}.

We apply the ranking procedure outline in section \ref{def-q} on two datasets. For target price rankings, we use the average per quarter daily errors within one quarter as the measure of analysts' forecasting ability (eq. \ref{tp:pmafe}).

Table \ref{tab:example} and figure \ref{fig:example} illustrate an example of how we estimate the \textit{PMAFE}. Four analysts had valid target price for Amazon for second quarter of 1999. We plot the daily Amazon price against the brokers' target prices. Table \ref{tab:example} shows the resulting TP and EPS ranking. On the bases of the average daily errors, LEGG is the most accurate in forecasting stock price and  DLJ is the least accurate. For the EPS case (panel B), PACCREST is the most accurate in EPS forecasting and RBRTSON is the least.

%The daily frequency allows us to capture the analysts' performance with better precision. Panel A of table (\ref{tab:example}) shows an example of rankings based on target prices. As we observe, during the second quarter of 1999, four analysts have valid target prices for Amazon (AMZN) stock. At the end of the quarter, we calculate the average daily error for each of the analysts and apply equations (\ref{tp:pmafe}) and (\ref{tp:rank}) to obtain the ranks of each of the analysts. Figure (\ref{fig:example}) demonstrates the daily AMZN stock price and the levels of the analysts' target prices. The model gives to brokers the follow ranking: LEGG(1), MONTSEC(2), KAUFBRO(3), DLJ(4). In this example, LEGG is the most accurate in forecasting the AMZN stock and DLJ is the least.
\subsection{Ranking contingency results}
We consider  three terciles (\textit{top}, \textit{medium}, \textit{bottom}). In one particular quarter ($t$), we place  analysts at one of these bins which corresponds to a tercile. We, then,  check analysts position at the immediate next quarter ($t+1$) and after one year ($t+4$).

We convert the rankings into scores as follows:
\begin{equation}
\label{eq:score}
score_{j,s}=\frac{rank_{j,s}}{\max{rank}}
\end{equation}

To get the cross-sectional values of scores across different stocks, we take the average of $score_{j,s}$
\begin{equation}
\label{eq:mean-score}
score_{j}= \frac{1}{k} \sum_{s=1}^{k} score_{j,s}
\end{equation}
where $k$ is number of stocks for which an analyst $j$ had rank.

Table (\ref{tab:rank-stat}) shows a contingency analysis of the ranks. 
%Chi square statistics include/ expected frequiency test
%Journ. of Fin economics 
Panel A shows the dynamics of each tercile for rankings based on the accuracy in target prices. We observe that analysts exhibit strong ranking consistency as, on average, they stay at the same tercile after one quarter and even after one year. The case of EPS ranking shows similar results (panel B). This analysis confirms recent finding of \cite{hilary2013} on analyst forecast consistency.

@@

Tables \ref{tab:view-stat} -- \ref{tab:view-stat-default} present the descriptive statistics of analysts' expected returns conditional on the information sets. The expected returns  are computed comparing TP estimates with actual price. To inform the strategies we obtain rank-weighted estimates where weights are given either by the TP or the EPS ranks.

@@

Panel A show the statistics for the consensus expectations as defined in eq. (\ref{consq}). The average consensus among the analysts about stock expected return for \tr{}, \naive{} and \default{} information sets are \Sexpr{round(stat.nr[year=='Total'&Method=='true',mean],3)*100 }\%, \Sexpr{round(stat.nr[year=='Total'&Method=='naive',mean],3)*100 }\%, and \Sexpr{round(stat.nr[year=='Total'&Method=='default',mean],3)*100 }\% respectively. Panel B of the table shows the TP accuracy weighted average expected returns (\Sexpr{ round(stat.pt.view[year=='Total'&Method=='true',mean],3)*100}\%, \Sexpr{ round(stat.pt.view[year=='Total'&Method=='naive',mean],3)*100}\%, and \Sexpr{ round(stat.pt.view[year=='Total'&Method=='default',mean],3)*100}\%). Panel C shows the EPS based weighted expected returns averages (\Sexpr{ round(stat.eps.view[year=='Total'&Method=='true',mean],3)*100}\%, \Sexpr{ round(stat.eps.view[year=='Total'&Method=='naive',mean],3)*100}\%, and \Sexpr{ round(stat.eps.view[year=='Total'&Method=='default',mean],3)*100}\%).

@@

\section{Empirical Results}
\label{sec:results}

We report results from trading strategies in table (\ref{tab:strategy}). We split the table into four panels: one for the passive strategy and three for each of the information availability sets.

>>>>

@@

<<chp1-load-data,echo=F,warning=FALSE,message=FALSE>>=
strategies <- c('true','recent','all-time')
setwd('~/Dropbox/workspace/Projects/BL-strategies')
library(reshape2)
library(descr)
library(data.table)
library(ggplot2)
library(stringr)
library(zoo)
library(xtable)
library(PerformanceAnalytics)
library(abind)
load('cache/q.data.RData')
load('cache/final.bl.RData')
load('cache/market.set.RData')
load('cache/example.dt.RData')
source('lib/aux.functions.R')
#library(ProjectTemplate)
#load.project()
###ADD NAs for the case of no ranking ( if NA, no raking)
trunk.percent <- 0.05
stocks <- unique(market.set[,Stock])
#q.data[,quantile(b.view,c(trunk.percent,1-trunk.percent),na.rm=T)]
q.data.full <- setkey(q.data[,':='(year=format(s.Date,'%Y'))],Stock)

@@

full.stat.call <- quote(list(min=min(N),mean=mean(N),median=as.character(median(N)),max=max(N),st.dev=sd(N)))

brok.full.stat.pt <- rbind(setkey(na.omit(core.dt)[,.N,by=.(Stock,q.id,year)][,eval(full.stat.call),by=year],year),setnames(cbind('Total',na.omit(core.dt)[,.N,by=.(Stock,q.id,year)][,eval(full.stat.call)]),1,'year'))

#Views stat
my.call=quote(list(mean=mean(View),median=median(View),st.dev=sd(View),Stocks=length(unique(Stock))))

stat.nr <- rbind(na.omit(nr.views)[,eval(my.call),by=.(year,Method)],setnames(cbind('Total',na.omit(nr.views)[,eval(my.call),by=.(Method)]),1,'year'))[,Strategy:='CONS']

stat.pt.view <- rbind(pt.rank.views[View!=0,eval(my.call),by=.(year,Method)],setnames(cbind('Total',pt.rank.views[View!=0,eval(my.call),by=.(Method)]),1,'year'))[,Strategy:='PT']

stat.eps.view <- rbind(eps.rank.views[View!=0,eval(my.call),by=.(year,Method)],setnames(cbind('Total',eps.rank.views[View!=0,eval(my.call),by=.(Method)]),1,'year'))[,Strategy:='EPS']

@@

brok.full.stat.eps <- rbind(setkey(na.omit(complete.dt)[,.N,by=.(Stock,q.id,year)][,eval(full.stat.call),by=year],year),setnames(cbind('Total',na.omit(complete.dt)[,.N,by=.(Stock,q.id,year)][,eval(full.stat.call)]),1,'year'))

@@

\chapter{Are rankings of financial analysts useful to investors?}
\section{Introduction}
\label{ch1-sec:intro}

@@

The paper is organized as follows: section \ref{ch1-sec:ranking} provides motivation to use rankings of financial analysts; section \ref{ch1-sec:trading} outlines our proposed  trading strategies; section \ref{ch1-sec:rankings} describes the sample and presents some preliminary results; section \ref{ch1-sec:results} presents and discusses the results; and section \ref{ch1-sec:conclusion} concludes.

\section{Industry Rankings of Financial Analysts}
\label{ch1-sec:ranking}

@@

%In addition, for each strategy, we have three timing scenarios of when we know the analysts' information: we use the \true{} timing for the case when the information is known at time $t$ (a hypothetical case); we use the \naive{} timing for the $t-1$ case, and we use the \default{} timing \citep{aiguzhinov2010} for the $t=1 \ldots t-1$ case.

@@

\section{Trading Strategies}
%- BL model
%- Information sets
\label{ch1-sec:trading}
Our trading strategy use the framework for  active portfolio management proposed by \cite{black1992}.  The model incorporates ``views" in a CAPM framework, forming optimal portfolios in a mean-variance optimization setting. ``Views" are expectations on individual stocks future performance by investors/analysts.

@@

We define a trading strategy as follows (figure \ref{ch1-fig:bl}):
\begin{enumerate}
\item  At the beginning of quarter $t$ for each stock $s$,   we define $Q$ and $\Omega$ (see \ref{ch1-def-q} and \ref{ch1-def-omega} below);

@@

\subsection{Defining $Q$}
\label{ch1-def-q}

For the consensus strategy, we use median of expected returns for a particular stock $s$:
\begin{equation}
\label{ch1-consq}
Q_{cons,s}= \median \left\{r_{j,s}\right\}
%\frac{1}{N} \sum_{j=1}^{N} r_{j,s}
\end{equation}
%$N$ is the number of analysts with a valid TP report and
where $r_{j,s}=TP_{j,s}/P_{s}-1$  is analyst's $j$ expected return computed using the analyst price target $TP_{j,s}$ and stock price $P_{s}$\footnote{Consistent with the literature, we use stock price 3 days \emph{ex-ante} the TP announcement. This is done to avoid any information leakage around new TP announcement day \citep{bonini2010}}.

@@

\begin{equation}
\label{ch1-eq:weight}
w_{j,s}=1-\frac{rank_{j,s}-\min{ \{rank \} }}{\max{\{rank \}}}
\end{equation}

The expected rank-weighted return is thus:
\begin{equation}
\label{ch1-rankq}
Q_{rank,s}=\frac{\sum_{j=1}^{N} (w_{j,s} \times r_{j,s})}{\sum_{j=1}^{N} w_{j,s}}
\end{equation}
$N$ is the number of analysts.

@@

\begin{equation}
\label{ch1-dfe}
FE_{i,s}^{TP}=|{P_{s}-TP_{j,s}}|
\end{equation}
The PMAFE is given as:
\begin{equation}
\label{ch1-tp:pmafe}
PMAFE_{j,s}^{TP}=\frac{FE_{i,s}^{TP}}{\overline{FE_{s}^{TP}}}
\end{equation}
where $\overline{{FE}_{t,s}^{TP}}$ is the average over quarter forecasting error. The target price is fixed over the quarter unless it gets revised.

The rank  that enters equation (\ref{ch1-eq:weight}) is average analyst's $PMAFE^{TP}$ per quarter:
\begin{equation}
\overline{PMAFE_{j,s}^{TP}}=\frac{1}{T} \sum_{t=1}^{T} PMAFE_{j,t,s}^{TP}
\end{equation}

\begin{equation}
\label{ch1-tp:rank}
rank_{j,s}=\rank_{j=1}^{N} \left\{ \overline{PMAFE_{j,s}^{TP}} \right\}
\end{equation}
where $T$ is the number of trading days in a quarter. Figure \ref{ch1-fig:example} shows an example.

\subsubsection{EPS ranking}
To compute the EPS rankings, we apply the same procedure as above:
\begin{equation}
FE_{j,s}^{EPS}=|{ACT_{s}-PRED_{j,s}}|
\end{equation}
\begin{equation}
PMAFE_{j,s}^{EPS}= \frac{FE_{j,s}^{EPS}}{\overline{FE_{s}^{EPS}}}
\end{equation}
\begin{equation}
\label{ch1-eps:rank}
rank_{j,s}=\rank_{j=1}^{N} \left\{ PMAFE_{j,s}^{EPS} \right\}
\end{equation}
where $ACT_{s}$ and $PRED_{j,s}$ are the actual and  analyst $j$ forecast EPS respectively for each stock $s$.

\subsection{Defining the confidence of expected returns $\Omega$}
\label{ch1-def-omega}
The confidence of $Q$ are given by the coefficient of variation of forecasting errors:

\begin{equation}
\label{ch1-omega}
CV = \frac{\sigma (FE)}{\overline{FE}}
\end{equation}
where $\sigma$ and $\overline{FE}$ are the standard deviation and the mean of the analysts' forecasting error for either TP or EPS. A low value of $CV$ reflects consensual estimates of either future prices or EPS.

@@

In addition to these sets,  we also create a hypothetical scenario that assumes we know the future information about the analyst' accuracy performance which would only be available at the end of $t$. In this setting, we are assuming that we could  perfectly predict the true analysts' ranking in advance. This represents the perfect foresight strategy; therefore, it serves as a reference point to the other trading strategies. We call this approach as the \true{} set.

We formalize these sets as follows.
\begin{itemize}
\item  the \true{} %-- a perfect foresight information:
\begin{equation}
\label{ch1-q:true}
\widehat{Q_{.,t}}=Q_{.,t}
\end{equation}

\item  the \naive{} % -- $t-$ information:
\begin{equation}
\label{ch1-q:naive}
\widehat{Q_{.,t}}=Q_{.,t-1}
\end{equation}

\item  the \default{} %-- the entire history of analysts performance:
%\begin{itemize}
%\item ranking based views:
\begin{equation}
\label{ch1-q:default}
\widehat{Q_{.,t}} = \frac{1}{T} \sum_{t=1}^{T} Q_{.,t}
\end{equation}
%\item consensus based views:
%\begin{equation}
%\label{ch1-default.cons}
%\widehat{consQ_{t,s}} = \frac{1}{T} \sum_{T=1}^{T-1} consQ_{t,s}
%\end{equation}
%\end{itemize}
\end{itemize}
where the $.$ subscript corresponds to either $Q_{cons,s}$ (eq. \ref{ch1-consq}) or $Q_{rank,s}$ (eq. \ref{ch1-rankq}).

\section{Data and preliminary results}
\label{ch1-sec:rankings}

@@

Consistent with other studies on analysts' expected returns that work with price targets  \citep{bradshaw2002,brav2003,da2011}, we truncate the sample of $TP/P-1$ at the \Sexpr{ trunk.percent *100}\textsuperscript{th} percentile (values below \Sexpr{round(q.data[,min(trunk.view,na.rm=T)],3)}) and at the \Sexpr{(1-trunk.percent)*100}\textsuperscript{th} percentile (values above \Sexpr{ round(q.data[,max(trunk.view,na.rm=T)],3)}). This is done due to occurrence of the extreme values. Most of these extreme values are driven by misalignment errors found on I/B/E/S data\footnote{We found some differences between the  DataStream and I/B/E/S the databases. In some cases the stock-splits and the dividends were not properly adjusted.}. After trimming, the number of observations ($\mathrm{Stock} \times \mathrm{Broker} \times  \mathrm{Quarter}$) is reduced  from \Sexpr{q.data[,.N]} to \Sexpr{q.data[!is.na(trunk.view),.N]}.

The initial file of quarterly EPS forecast consists of  \Sexpr{eps.dt[,.N,by=Broker][,.N]} brokers covering \Sexpr{eps.dt[,.N,by=Stock][,.N]} stocks. To implement the EPS ranking, we require that a stock had at least three brokers per quarter. In addition, we require that a broker has to be active in covering a particular stock for at least 3 years (12 quarters). Our final sample of EPS forecasts consist of  \Sexpr{complete.dt[,.N,by=Broker][,.N]} brokers covering \Sexpr{complete.dt[,.N,by=Stock][,.N]} stocks. The total number of observation is \Sexpr{complete.dt[,.N]}.

Table (\ref{ch1-tab:ret-stat}) shows the distribution of the final sample of target price and EPS data. Panel A shows the number of quarterly target prices per stock. For the sample period (1999-2009), we report \Sexpr{core.dt[,.N,by=list(Stock,q.id)][,.N]} stock-quarter observations. Each stock had on average of \Sexpr{round(core.dt[,.N,by=list(Stock,q.id)][,mean(N)],2)} quarterly target price reports. Panel B  presents similar statistics of the number of EPS forecasts. In total, the EPS forecast sample has \Sexpr{complete.dt[,.N,by=list(Stock,q.id)][,.N]} stock-quarter observations. The average number of quarterly forecasts is \Sexpr{round(complete.dt[,.N,by=list(Stock,q.id)][,mean(N)],2)}.

We apply the ranking procedure outline in section \ref{ch1-def-q} on two datasets. For target price rankings, we use the average per quarter daily errors within one quarter as the measure of analysts' forecasting ability (eq. \ref{ch1-tp:pmafe}).

Table \ref{ch1-tab:example} and figure \ref{ch1-fig:example} illustrate an example of how we estimate the \textit{PMAFE}. Four analysts had valid target price for Amazon for second quarter of 1999. We plot the daily Amazon price against the brokers' target prices. Table \ref{ch1-tab:example} shows the resulting TP and EPS ranking. On the bases of the average daily errors, LEGG is the most accurate in forecasting stock price and  DLJ is the least accurate. For the EPS case (panel B), PACCREST is the most accurate in EPS forecasting and RBRTSON is the least.

%The daily frequency allows us to capture the analysts' performance with better precision. Panel A of table (\ref{ch1-tab:example}) shows an example of rankings based on target prices. As we observe, during the second quarter of 1999, four analysts have valid target prices for Amazon (AMZN) stock. At the end of the quarter, we calculate the average daily error for each of the analysts and apply equations (\ref{ch1-tp:pmafe}) and (\ref{ch1-tp:rank}) to obtain the ranks of each of the analysts. Figure (\ref{ch1-fig:example}) demonstrates the daily AMZN stock price and the levels of the analysts' target prices. The model gives to brokers the follow ranking: LEGG(1), MONTSEC(2), KAUFBRO(3), DLJ(4). In this example, LEGG is the most accurate in forecasting the AMZN stock and DLJ is the least.
\subsection{Ranking contingency results}
We consider  three terciles (\textit{top}, \textit{medium}, \textit{bottom}). In one particular quarter ($t$), we place  analysts at one of these bins which corresponds to a tercile. We, then,  check analysts position at the immediate next quarter ($t+1$) and after one year ($t+4$).

We convert the rankings into scores as follows:
\begin{equation}
\label{ch1-eq:score}
score_{j,s}=\frac{rank_{j,s}}{\max{rank}}
\end{equation}

To get the cross-sectional values of scores across different stocks, we take the average of $score_{j,s}$
\begin{equation}
\label{ch1-eq:mean-score}
score_{j}= \frac{1}{k} \sum_{s=1}^{k} score_{j,s}
\end{equation}
where $k$ is number of stocks for which an analyst $j$ had rank.

Table (\ref{ch1-tab:rank-stat}) shows a contingency analysis of the ranks.
%Chi square statistics include/ expected frequiency test
%Journ. of Fin economics
Panel A shows the dynamics of each tercile for rankings based on the accuracy in target prices. We observe that analysts exhibit strong ranking consistency as, on average, they stay at the same tercile after one quarter and even after one year. The case of EPS ranking shows similar results (panel B). This analysis confirms recent finding of \cite{hilary2013} on analyst forecast consistency.

@@

Tables \ref{ch1-tab:view-stat} -- \ref{ch1-tab:view-stat-default} present the descriptive statistics of analysts' expected returns conditional on the information sets. The expected returns  are computed comparing TP estimates with actual price. To inform the strategies we obtain rank-weighted estimates where weights are given either by the TP or the EPS ranks.

@@

Panel A show the statistics for the consensus expectations as defined in eq. (\ref{ch1-consq}). The average consensus among the analysts about stock expected return for \true{}, \naive{} and \default{} information sets are \Sexpr{round(stat.nr[year=='Total'&Method=='true',mean],3)*100 }\%, \Sexpr{round(stat.nr[year=='Total'&Method=='naive',mean],3)*100 }\%, and \Sexpr{round(stat.nr[year=='Total'&Method=='default',mean],3)*100 }\% respectively. Panel B of the table shows the TP accuracy weighted average expected returns (\Sexpr{ round(stat.pt.view[year=='Total'&Method=='true',mean],3)*100}\%, \Sexpr{ round(stat.pt.view[year=='Total'&Method=='naive',mean],3)*100}\%, and \Sexpr{ round(stat.pt.view[year=='Total'&Method=='default',mean],3)*100}\%). Panel C shows the EPS based weighted expected returns averages (\Sexpr{ round(stat.eps.view[year=='Total'&Method=='true',mean],3)*100}\%, \Sexpr{ round(stat.eps.view[year=='Total'&Method=='naive',mean],3)*100}\%, and \Sexpr{ round(stat.eps.view[year=='Total'&Method=='default',mean],3)*100}\%).

@@

\section{Empirical Results}
\label{ch1-sec:results}

We report results from trading strategies in table (\ref{ch1-tab:strategy}). We split the table into four panels: one for the passive strategy and three for each of the information availability sets.
